---
title: "ddPCR"
author: "Emily Lacroix"
date: "Last update: 1/14/2022"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set-up

## Libraries
```{r warning=FALSE, message=FALSE}

library(readxl)
library(MASS)
library(car)
library(scales)
library(tidyverse)
library(conflicted)

conflicted::conflict_prefer("select", "dplyr")
conflicted::conflict_prefer("filter", "dplyr")


```

## Files

```{r eval=FALSE}
data_file <- "StanfordDish_AllData.xlsx"
```


```{r include=FALSE}

data_file <- "~/Box Sync/Stanford (elacroix@stanford.edu)/Research/StanfordDish/Data/StanfordDish_AllData.xlsx"

```

## Labels 
```{r}
site_labels <- c(
  "DISHC" = "Clay\nLoam",
  "DISHT" = "Loam",
  "DISHR" = "Sandy\n Loam"
)
```

## Figure theme
```{r}
my_theme <- function(base_size = 12, base_family = ""){ ## Control base font face and size. use `rel()` for relative font size.
  theme_bw(base_size = base_size, base_family = base_family) %+replace%
    theme(
      panel.border = element_rect(colour = "black", fill = "transparent"),
      panel.background  = element_blank(),
      panel.grid = element_blank(),
      strip.background = element_blank(),
      legend.position="top",
      legend.title = element_blank(),
      strip.text = element_text(size = 13)
    )
}

```


## Load data
```{r}

ddpcr <- 
  read_xlsx(data_file, sheet = "ddPCR") %>% 
  mutate(
    across(
      site, 
      ~factor(., levels = c("DISHC", "DISHT", "DISHR"), labels = site_labels)
    )
  )

```


# 16S - absolute abundance

## Figure S7
```{r}
ddpcr %>% 
  group_by(site) %>% 
  summarise(
    mean_16s = mean(copies_per_g_16s),
    se_16s = sd(copies_per_g_16s) / sqrt(n())
  ) %>% 
  ggplot(aes(x = site, y = mean_16s)) + 
  geom_pointrange(
    aes(ymin = mean_16s - se_16s, ymax = mean_16s + se_16s),
    fatten = 4,
    size = 1
  ) +
  my_theme() + 
  theme(
    axis.text = element_text(size = 18),
    axis.title = element_text(size = 20),
    aspect.ratio = 1.5
  ) +
  labs(
    y = "16S rRNA (copies per g dry soil) ",
    x = NULL
  )

```

# mcrA - absolute abundance

## Figure S8
```{r}
ddpcr %>% 
  group_by(site) %>% 
  summarise(
    mean_mcrA = mean(copies_per_g_mcra, na.rm = TRUE),
    se_mcrA = sd(copies_per_g_mcra, na.rm = TRUE) / sqrt(n())
  ) %>% 
  ggplot(aes(x = site, y = mean_mcrA)) + 
  geom_pointrange(
    aes(ymin = mean_mcrA - se_mcrA, ymax = mean_mcrA + se_mcrA),
    fatten = 4,
    size = 1
  ) +
  scale_y_continuous(labels = scientific) +
  my_theme() + 
  theme(
    axis.text = element_text(size = 18),
    axis.title = element_text(size = 20),
    aspect.ratio = 1.5
  ) +
  labs(
    y = "mcrA (copies per g dry soil) ",
    x = NULL
  )
```

## Statistics 

Non-transformed mcrA copies per g are normally distributed with very unequal variances.

### Check normality of log transformed copies per g mcrA

```{r}

ddpcr %>% 
  mutate(log_copies = log(copies_per_g_mcra + 5656)) %>% #adding half the lowest non-zero value
  group_by(site) %>% 
  summarize(p_normality = shapiro.test(log_copies)$p.value) 

```

### Check for equal variances of log transformed copies per g mcrA

```{r}

leveneTest(
  log_copies ~ site, 
  data = 
    ddpcr %>% mutate(log_copies = log(copies_per_g_mcra + 5656))
)

```
Log transformed variances ARE equal. 

### ANOVA of log transformed copies per g mcrA

```{r}
aov_conc_site_manual <- 
  aov(
  log_copies ~ site, 
    ddpcr %>% mutate(log_copies = log(copies_per_g_mcra + 5656))
)

aov_conc_site_manual %>% summary()

TukeyHSD(aov_conc_site_manual) 

```

There are significant differences between:
1) Loam & Clayey Loam 
2) Sandy Loam & Clayey Loam 


# Normalized mcrA

## Figure 5

```{r}
ddpcr %>% 
  group_by(site) %>% 
  summarize(
    mean_prop_mcra = mean(proportion_mcra),
    se_prop_mcra = sd(proportion_mcra) / sqrt(n())
  ) %>% 
  ggplot(aes(x = site, y = mean_prop_mcra)) + 
  geom_pointrange(
    aes(
      ymin = mean_prop_mcra - se_prop_mcra, 
      ymax = mean_prop_mcra + se_prop_mcra
    ),
    fatten = 4,
    size = 1
  ) + 
  my_theme() + 
  theme(
    axis.text = element_text(size = 18),
    axis.title = element_text(size = 20),
    aspect.ratio = 1.5
  ) +
  labs(
    y = "copies mcrA / copies 16S",
    x = NULL
  )


```


## Statistics 

### Check normality 
```{r}

ddpcr %>%   
  group_by(site) %>% 
  summarize(p_normality = shapiro.test(proportion_mcra)$p.value)
  
```

### Check for equal variances 

```{r}
leveneTest(
  proportion_mcra ~ site, 
  data = ddpcr
)
```


Proportion mcrA is normally distributed with equal variances. 

### ANOVA
```{r}

aov_prop_mcra <- 
  aov(
  proportion_mcra ~ site, 
  data = ddpcr
)

aov_prop_mcra %>% summary()

TukeyHSD(aov_prop_mcra) 

```

Like the absolute abundance mcrA data, there is a significant difference between:
1) Loam & Clayey Loam
2) Sandy Loam & Clayey Loam

