---
title: "MAOC_POC"
author: "Emily Lacroix"
date: "Last update: 1/19/2022"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script analyzes the Density Fractionation data from the soils used in both the Disturbance and Oxygen Enrichment Incubations.

# Setup

## Load libraries
```{r}

library(readxl)
library(car)
library(tidyverse)

options(dplyr.summarise.inform = FALSE)
```

## Files

```{r eval=FALSE}
data_file <- "StanfordDish_AllData.xlsx"
```

```{r include=FALSE}
data_file <- "~/Box Sync/Stanford (elacroix@stanford.edu)/Research/StanfordDish/Data/StanfordDish_AllData.xlsx"
```

## Constants 
```{r}
control_cores <- c(1, 5, 9, 13)

```

## Figure theme
```{r}
my_theme <- function(base_size = 12, base_family = ""){ ## Control base font face and size. use `rel()` for lelative font size.
  theme_bw(base_size = base_size, base_family = base_family) %+replace%
    theme(
      panel.border = element_rect(colour = "black", fill = "transparent"),
      panel.background  = element_blank(),
      panel.grid = element_blank(),
      strip.background = element_blank(),
      legend.position="top",
      legend.title = element_blank(),
      strip.text = element_text(size = 13)
    )
}

```

## Figure labels
```{r}
site_labels <- c(
  "DISHC" = "Clayey\nLoam",
  "DISHT" = "Loam",
  "DISHR" = "Sandy\n Loam"
)

```


# Import and Combine Data 

## Import data
```{r}
dens_frac <- 
  data_file %>% 
  read_xlsx(sheet = "DensityFractionation")

tc_tn <-
  data_file %>% 
  read_xlsx(sheet = "TotalCarbon")
```

## Combine Data
```{r}

maoc_poc <-
  dens_frac %>% 
  left_join(
    tc_tn %>% 
      group_by(site) %>% 
      summarise(avg_perc_c = mean(perc_c, na.rm = TRUE)),
    by = "site"
  ) %>% 
  mutate(
    estimated_total_mg_c = 1000 * sample_mass * (avg_perc_c / 100),
    prop_maoc = mg_c_hf / estimated_total_mg_c,
    prop_poc = mg_c_lf / estimated_total_mg_c,
    across("site", factor, levels = c("DISHC", "DISHT", "DISHR")),
    percent_recovery = 100* (total_g_lf + mass_hf) / sample_mass
  ) 


```

# Figures, Tables, Calculations

##  Percent Recovery

```{r}
maoc_poc %>% pull(percent_recovery) %>% summary()
```

## Table 1: MAOC & POC 

```{r}
maoc_poc %>% 
  group_by(site) %>% 
  summarise(
    avg_perc_poc = 100 * mean(prop_poc),
    se_perc_poc = 100 * sd(prop_poc) / sqrt(n()),
    avg_perc_maoc = 100 * mean(prop_maoc),
    se_perc_maoc = 100 * sd(prop_maoc) / sqrt(n())
  ) 
```

# Statistics

## POC

### Check for normality
```{r}
maoc_poc %>% 
  group_by(site) %>% 
  summarise(
    normality_stat = shapiro.test(prop_poc)$statistic,
    normality_p_value = shapiro.test(prop_poc)$p.value
  )

```


### Check for equal variances

```{r}
leveneTest(
  prop_poc ~ site, 
  data = maoc_poc 
)

```
The sites proportion of POC are normally distributed with equal variances. 

### ANOVA

```{r}
summary(aov(prop_poc ~ site, data = maoc_poc)) 

```

They are significantly different! 
Now do Tukey's HSD

```{r}
TukeyHSD(aov(prop_poc ~ site, data = maoc_poc))
```

Only DISHR is different from DISHC, otherwise there is slight overlap it appears

## MAOC

### Check for normality
```{r}
maoc_poc %>% 
  group_by(site) %>% 
  summarise(
    normality_stat = shapiro.test(prop_maoc)$statistic,
    normality_p_value = shapiro.test(prop_maoc)$p.value
  )


```

### Check for equal variances
```{r}
leveneTest(
  prop_maoc ~ site, 
  data = maoc_poc 
)
```

Proportion MAOC is also normally distributed with equal variances. 

### ANOVA

```{r}
summary(aov(prop_maoc ~ site, data = maoc_poc)) 
```

