---
title: "FeMn"
author: "Emily Lacroix"
date: "Last update: 1/14/2022"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

## Load libraries
```{r}
library(tidyverse)
library(readxl)
library(multcompView)
library(car)

options(dplyr.summarise.inform = FALSE)
```

## Files

```{r eval=FALSE}
data_file <- "StanfordDish_AllData.xlsx"
```

```{r include=FALSE}
data_file <- "~/Box Sync/Stanford (elacroix@stanford.edu)/Research/StanfordDish/Data/StanfordDish_AllData.xlsx"
```

## Constants 
```{r}
control_cores <- c(1, 5, 9, 13)

```

## Figure theme
```{r}
my_theme <- function(base_size = 12, base_family = ""){ ## Control base font face and size. use `rel()` for relative font size.
  theme_bw(base_size = base_size, base_family = base_family) %+replace%
    theme(
      panel.border = element_rect(colour = "black", fill = "transparent"),
      panel.background  = element_blank(),
      panel.grid = element_blank(),
      strip.background = element_blank(),
      legend.position="top",
      legend.title = element_blank(),
      strip.text = element_text(size = 13)
    )
}

```

## Figure labels
```{r}
site_labels <- c(
  "DISHC" = "Clay\nLoam",
  "DISHT" = "Loam",
  "DISHR" = "Sandy\nLoam"
)

```

## Useful Functions
```{r}
tri.to.squ<-function(x)
{
  rn <- row.names(x)
  cn <- colnames(x)
  an <- unique(c(cn,rn))
  myval <- x[!is.na(x)]
  mymat <- 
    matrix(1, nrow = length(an), ncol = length(an), dimnames = list(an,an))
  for(ext in 1:length(cn))
  {
    for(int in 1:length(rn))
    {
      if(is.na(x[row.names(x) == rn[int], colnames(x) == cn[ext]])) next
      mymat[row.names(mymat) == rn[int], colnames(mymat) == cn[ext]] <- 
        x[row.names(x) == rn[int], colnames(x) == cn[ext]]
      mymat[row.names(mymat) == cn[ext], colnames(mymat) == rn[int]] <- 
        x[row.names(x) == rn[int], colnames(x) == cn[ext]]
    }
  }
  return(mymat)
}
```

# Import and Combine Data 

## Import data

```{r}

fe_mn <- 
  data_file %>% 
  read_xlsx(sheet = "FeMnHCl")

dist_grav <- 
  data_file %>% 
  read_xlsx(sheet = "GravimetricData - Oxic")

```

## Combine

This code:
* joins the Fe and Mn data with the associated moisture data for the core
* Categorizes cores as disturbed or undisturbed
* Assigns the cores to a WFPS bin
* Eliminates cores with < 15% WFPS 
* Makes sites and WFPS category into factors 

```{r}

fe_mn_data <-
  fe_mn %>% 
  full_join(dist_grav, by = c("site", "core", "experiment")) %>% 
  mutate(
    wfps_bin = 
      factor(
        case_when(
          wfps <= 60 ~ "15 - 60 %",
          is.na(wfps) ~ "15 - 60 %",
          experiment == "OxEn" ~ "15 - 60 %",
          wfps > 60 ~ "> 60 %",
          TRUE ~ NA_character_
        ),
        levels = c("15 - 60 %", "> 60 %")
      )
  ) %>% 
  filter(wfps >= 15 | experiment == "OxEn") %>% 
  mutate(
    control_disturbed = 
      factor(
        if_else(
          (core %in% control_cores | experiment == "OxEn"), 
          "Undisturbed", 
          "Disturbed"
        ),
        levels = c("Undisturbed", "Disturbed")
      ),
    across("site", factor, levels = c("DISHC", "DISHT", "DISHR"), labels = site_labels),
    across(wfps_bin, factor, levels = c("15 - 60 %", "> 60 %"))
  ) 

fe_mn

```


# Figures and Analysis

## Fe II to III 

```{r}

fe_ii_to_iii_averages <- 
  fe_mn_data %>% 
  filter(wfps_bin == "15 - 60 %") %>% 
  group_by(site) %>% 
  summarise(
    site_avg_fe_ii_to_iii = mean(fe_ii_to_iii, na.rm = TRUE),
    site_se_fe_ii_to_iii = sd(fe_ii_to_iii, na.rm = TRUE) / sqrt(n())
  )

fe_ii_to_iii_averages

```
### Figure 4

```{r}
fe_mn_data %>% 
  group_by(site, wfps_bin, control_disturbed) %>% 
  summarise(
    avg_fe_ii_to_iii = mean(fe_ii_to_iii, na.rm = TRUE),
    se_fe_ii_to_iii = sd(fe_ii_to_iii, na.rm = TRUE) / sqrt(n())
  ) %>% 
  left_join(fe_ii_to_iii_averages, by = "site") %>% 
  filter(wfps_bin == "15 - 60 %") %>% 
  ggplot(aes(x = site, group = control_disturbed)) +
  geom_col(aes(y = avg_fe_ii_to_iii, fill = control_disturbed), position = "dodge") +
  geom_errorbar(
    aes(ymin = avg_fe_ii_to_iii, ymax = avg_fe_ii_to_iii + se_fe_ii_to_iii),
    position = "dodge",
  ) +
  geom_errorbar(
    aes(
      ymin = site_avg_fe_ii_to_iii - site_se_fe_ii_to_iii,
      ymax = site_avg_fe_ii_to_iii + site_se_fe_ii_to_iii
    ),
    color = "gold",
    width = 0,
    size = 1.5
  ) + 
  geom_point(
    aes(
      y = site_avg_fe_ii_to_iii, 

    ), 
    color = "gold", 
    shape = 18, 
    size = 6
  ) + 
  scale_fill_manual(values = c("black", "darkred")) + 
  my_theme() +
  theme(
    aspect.ratio = 1.5,
    legend.position = "right",
    axis.text = element_text(size = 16),
    axis.title = element_text(size = 18),
    legend.text = element_text(size = 16)
  ) +
  labs(
    x = NULL,
    y = expression(paste(Fe^`2+`~`/`~Fe[tot]))
  )

```

### Table S1: Fe/Mn replicates

```{r}
fe_mn_data %>% 
  group_by(site, wfps_bin, control_disturbed) %>% 
  summarise(
    avg_fe_ii_to_iii = mean(fe_ii_to_iii, na.rm = TRUE),
    se_fe_ii_to_iii = sd(fe_ii_to_iii, na.rm = TRUE) / sqrt(n()),
    reps = n()
  ) %>% 
  filter(wfps_bin == "15 - 60 %")
```


### Statistics - by bin

#### Check normality 
```{r}
fe_mn_data %>% 
  filter(wfps_bin == "15 - 60 %") %>% 
  group_by(site, wfps_bin, control_disturbed) %>%
  summarise(
    normality_stat = shapiro.test(fe_ii_to_iii)$statistic,
    normality_p_value = shapiro.test(fe_ii_to_iii)$p.value
  ) #common transformations do not help normality


```
#### Check for equal variances

```{r}
leveneTest(
  fe_ii_to_iii ~ site*wfps_bin*control_disturbed, 
  data = fe_mn_data %>% filter(wfps_bin == "15 - 60 %")
)
```


Non-normally distributed data with unequal variances. Pairwise wilcoxon rank sum test would be appropriate

#### Pairwise wilcoxon rank sum

```{r warning=FALSE}
wilcoxon_fe_mn_data <-
  fe_mn_data %>% 
  mutate(
    wfps_id = 
      case_when(
        wfps_bin == "15 - 60 %" ~ "dry",
        wfps_bin == "> 60 %" ~ "wet",        
      ),
    wilcoxon_id = str_c(site,"_",control_disturbed,"_",wfps_id)
  )

fe_ii_to_iii_wilcox <- 
  pairwise.wilcox.test(
  x = wilcoxon_fe_mn_data$fe_ii_to_iii,
  g = wilcoxon_fe_mn_data$wilcoxon_id,
  p.adjust.method = "BH",
  paired = FALSE
)

fe_ii_to_iii_matrix <- tri.to.squ(fe_ii_to_iii_wilcox$p.value)
fe_ii_to_iii_letters <- 
  multcompLetters(
    fe_ii_to_iii_matrix, 
    compare ="<=", 
    threshold = 0.05, 
    Letters = letters
  )

fe_ii_to_iii_letters
```


### Statistics - by site

```{r}
fe_mn_data %>% 
  filter(wfps_bin == "15 - 60 %") %>% 
  group_by(site) %>%
  summarise(
    normality_stat = shapiro.test(fe_ii_to_iii)$statistic,
    normality_p_value = shapiro.test(fe_ii_to_iii)$p.value
  ) #common transformations do not help normality

fe_mn_data %>% 
  filter(wfps_bin == "15 - 60 %") %>% 
  ggplot() +
  geom_histogram(aes(x = log(fe_ii_to_iii), fill = site), alpha = 0.2)


leveneTest(
  fe_ii_to_iii ~ site, 
  data = fe_mn_data %>% filter(wfps_bin == "15 - 60 %")
)
```

Non-normally distributed data with unequal variances. Pairwise wilcoxon rank sum test would be appropriate

#### Pairwise Wilcoxon by site 
```{r warning=FALSE}

wilcoxon_fe_mn_data_site <-
  fe_mn_data %>% 
  filter(wfps_bin == "15 - 60 %")

fe_ii_to_iii_wilcox_site <- 
  pairwise.wilcox.test(
  x = wilcoxon_fe_mn_data_site$fe_ii_to_iii,
  g = wilcoxon_fe_mn_data_site$site,
  p.adjust.method = "BH",
  paired = FALSE
)

fe_ii_to_iii_matrix_site <- tri.to.squ(fe_ii_to_iii_wilcox_site$p.value)
fe_ii_to_iii_letters_site <- 
  multcompLetters(
    fe_ii_to_iii_matrix_site, 
    compare ="<=", 
    threshold = 0.05, 
    Letters = letters
  )

fe_ii_to_iii_letters_site

fe_ii_to_iii_wilcox_site
```




## Fe2+ 

### Figure S6
```{r}
fe_ii_averages <- 
  fe_mn_data %>% 
  filter(wfps_bin == "15 - 60 %") %>% 
  group_by(site) %>% 
  summarise(
    site_avg_fe_ii = mean(fe_ii_mg_kg, na.rm = TRUE),
    site_se_fe_ii = sd(fe_ii_mg_kg, na.rm = TRUE) / sqrt(n())
  )

fe_ii_averages
```



```{r}
fe_mn_data %>% 
  filter(wfps_bin == "15 - 60 %") %>% 
  group_by(site, wfps_bin, control_disturbed) %>% 
  summarise(
    avg_fe_ii = mean(fe_ii_mg_kg, na.rm = TRUE),
    se_fe_ii = sd(fe_ii_mg_kg, na.rm = TRUE) / sqrt(n())
  ) %>% 
  left_join(fe_ii_averages, by = "site") %>% 
  ggplot(aes(x = site, fill = control_disturbed)) +
  geom_col(aes(y = avg_fe_ii), position = "dodge") + 
  geom_errorbar(
    aes(ymin = avg_fe_ii, ymax = avg_fe_ii + se_fe_ii),
    position = "dodge"
  ) + 
  geom_errorbar(
    aes(
      ymin = site_avg_fe_ii - site_se_fe_ii,
      ymax = site_avg_fe_ii + site_se_fe_ii
    ),
    color = "gold",
    width = 0,
    size = 1.5
  ) + 
  geom_point(
    aes(
      y = site_avg_fe_ii, 

    ), 
    color = "gold", 
    shape = 18, 
    size = 6,
    show.legend = FALSE
  ) + 
  scale_fill_manual(values = c("black", "darkred")) + 
  my_theme() + 
  theme(
    aspect.ratio = 1.5,
    legend.position = "right",
    axis.text = element_text(size = 16),
    axis.title = element_text(size = 18),
    legend.text = element_text(size = 16)
  ) +
  labs(
    x = NULL,
    y = expression(paste(mg~Fe^`2+`~kg^`-1`~soil))
  )

```

### Statistics - by bin

#### Check for normality
```{r}
fe_mn_data %>% 
  filter(wfps_bin == "15 - 60 %") %>% 
  group_by(site, wfps_bin, control_disturbed) %>%
  summarise(
    normality_stat = shapiro.test(log(fe_ii_mg_kg))$statistic,
    normality_p_value = shapiro.test(log(fe_ii_mg_kg))$p.value
  ) #common transformations do not help normality

fe_mn_data %>% 
  ggplot(aes(x = fe_ii_mg_kg)) + 
  geom_histogram() + 
  facet_grid(cols = vars(site), rows = vars(wfps_bin))


```
Non-normal data. Transformations (not shown) do not seem to help.

#### Check for equal variances
```{r}
leveneTest(
  fe_ii_mg_kg ~ site*wfps_bin*control_disturbed, 
  data = fe_mn_data 
)
```
Unequal variances.

#### Pairwise wilcoxon rank sum 
```{r warning=FALSE}
fe_ii_wilcox <- 
  pairwise.wilcox.test(
  x = wilcoxon_fe_mn_data$fe_ii_mg_kg,
  g = wilcoxon_fe_mn_data$wilcoxon_id,
  p.adjust.method = "BH",
  paired = FALSE
)

fe_ii_matrix <- tri.to.squ(fe_ii_wilcox$p.value)
fe_ii_letters <- multcompLetters(fe_ii_matrix, compare ="<=", threshold = 0.05, Letters = letters)

fe_ii_letters
```


### Statistics - by site

#### Check for normality 
```{r}
fe_mn_data %>% 
  filter(wfps_bin == "15 - 60 %") %>% 
  group_by(site) %>%
  summarise(
    normality_stat = shapiro.test(fe_ii_mg_kg)$statistic,
    normality_p_value = shapiro.test(fe_ii_mg_kg)$p.value
  ) #common transformations do not help normality


fe_mn_data %>% 
  ggplot(aes(x = fe_ii_mg_kg)) + 
  geom_histogram() + 
  facet_grid(cols = vars(site), rows = vars(wfps_bin))

```

#### Check for equal variances
```{r}
leveneTest(
  fe_ii_mg_kg ~ site*wfps_bin*control_disturbed, 
  data = fe_mn_data 
)
```


#### Pairwise Wilcoxon rank sum
```{r}
fe_ii_wilcox <- 
  pairwise.wilcox.test(
  x = wilcoxon_fe_mn_data$fe_ii_mg_kg,
  g = wilcoxon_fe_mn_data$site,
  p.adjust.method = "BH",
  paired = FALSE
)

fe_ii_matrix_site <- tri.to.squ(fe_ii_wilcox$p.value)
fe_ii_letters_site <- multcompLetters(fe_ii_matrix_site, compare ="<=", threshold = 0.05, Letters = letters)

fe_ii_letters_site
```


## Mn 

### Figure S5
```{r}
mn_averages <- 
  fe_mn_data %>% 
  filter(wfps_bin == "15 - 60 %") %>% 
  group_by(site) %>% 
  summarise(
    site_avg_mn = mean(mn_mg_kg, na.rm = TRUE),
    site_se_mn = sd(mn_mg_kg, na.rm = TRUE) / sqrt(n())
  )


```


```{r}
fe_mn_data %>% 
  filter(wfps_bin == "15 - 60 %") %>% 
  group_by(site, wfps_bin, control_disturbed) %>% 
  summarise(
    avg_mn = mean(mn_mg_kg, na.rm = TRUE),
    se_mn = sd(mn_mg_kg, na.rm = TRUE) / sqrt(n())
  ) %>% 
  left_join(mn_averages, by = "site") %>% 
  ggplot(aes(x = site, fill = control_disturbed)) +
  geom_col(aes(y = avg_mn), position = "dodge") + 
  geom_errorbar(
    aes(ymin = avg_mn, ymax = avg_mn + se_mn),
    position = "dodge"
  ) + 
  geom_errorbar(
    aes(
      ymin = site_avg_mn - site_se_mn,
      ymax = site_avg_mn + site_se_mn
    ),
    color = "gold",
    width = 0,
    size = 1.5
  ) + 
  geom_point(
    aes(
      y = site_avg_mn, 

    ), 
    color = "gold", 
    shape = 18, 
    size = 6,
    show.legend = FALSE
  ) + 
  scale_fill_manual(values = c("black", "darkred")) + 
  my_theme() + 
  theme(
    aspect.ratio = 1.5,
    legend.position = "right",
    axis.text = element_text(size = 16),
    axis.title = element_text(size = 18),
    legend.text = element_text(size = 16)
  ) +
  labs(
    x = NULL,
    y = expression(paste(mg~extractable~Mn~kg^`-1`~soil))
  )

```

### Statistics - by bin

#### Check for normality
```{r}

fe_mn_data %>% 
  filter(wfps_bin == "15 - 60 %") %>% 
  group_by(site, wfps_bin, control_disturbed) %>%
  summarise(
    normality_stat = shapiro.test(mn_mg_kg)$statistic,
    normality_p_value = shapiro.test(mn_mg_kg)$p.value
  )


fe_mn_data %>% 
  filter(wfps_bin == "15 - 60 %") %>% 
  ggplot(aes(x = mn_mg_kg)) + 
  geom_histogram() + 
  facet_grid(cols = vars(site), rows = vars(wfps_bin))



```

#### Check for equal variances

```{r}
leveneTest(
  mn_mg_kg ~ site*wfps_bin*control_disturbed, 
  data = fe_mn_data 
)

```


They are normally distributed with unequal variances. Will need to use Welch's ANOVA! 

#### Welch's ANOVA
```{r}
oneway.test(
  mn_mg_kg ~ site*control_disturbed, 
  data = fe_mn_data %>% filter(wfps_bin == "15 - 60 %"), 
  var.equal = FALSE
)

mn_aov <- 
  aov(  
    mn_mg_kg ~ site*control_disturbed, 
    data = fe_mn_data %>% filter(wfps_bin == "15 - 60 %"), 
    var.equal = FALSE
  )

```
Now I can do a Tukey's HSD as a follow-up

```{r}
TukeyHSD(mn_aov)
```
The only significant differences are between the sites. Going to simplify analysis to focus solely on site-wide differences


### Statistics- by site

#### Check normality

NOTE: a transformation was necessary to achieve normal data
```{r}

fe_mn_data %>% 
  filter(wfps_bin == "15 - 60 %") %>% 
  group_by(site) %>%
  summarise(
    normality_stat = shapiro.test((mn_mg_kg)^0.33333333)$statistic,
    normality_p_value = shapiro.test((mn_mg_kg)^0.33333333)$p.value
  )
 #will do stats on transformed values

```


#### Check for equal variances
```{r}
leveneTest(
  (mn_mg_kg)^0.33333333 ~ site, 
  data = fe_mn_data 
)

```

Will transform mn data by taking cube root and then performing ANOVA.

#### ANOVA on transformed (cube root) data
```{r}

mn_stats_data <- 
  fe_mn_data %>% 
  filter(wfps_bin == "15 - 60 %") %>% 
  mutate(
    mn_transformed = mn_mg_kg ^ (1/3)
  )

mn_aov <- aov(mn_transformed ~ site, data = mn_stats_data)

summary(mn_aov)

TukeyHSD(mn_aov)
```





