---
title: "Incubation GHG Fluxes"
author: "Emily Lacroix"
date: "Last update 1/19/2022"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

## Load libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(multcompView)
library(car)

options(dplyr.summarise.inform = FALSE)
```

## Files
```{r eval=FALSE}
data_file <- "StanfordDish_AllData.xlsx"
```


```{r include=FALSE}
data_file <- "~/Box Sync/Stanford (elacroix@stanford.edu)/Research/StanfordDish/Data/StanfordDish_AllData.xlsx"
```

## Constants 
```{r}
control_cores <- c(1, 5, 9, 13)

soil_core_area_m2 <- 0.00203

anoxic_soil_core_area_m2 <- 0.0005067

s_in_hr <- 3600

hr_in_day <- 24

umol_to_mmol <- 0.001

anoxic_incubation_period_hr <- 3

mols_per_l <- 1 / 22.4


```

## Figure theme
```{r}
my_theme <- function(base_size = 12, base_family = ""){ ## Control base font face and size. use `rel()` for lelative font size.
  theme_bw(base_size = base_size, base_family = base_family) %+replace%
    theme(
      panel.border = element_rect(colour = "black", fill = "transparent"),
      panel.background  = element_blank(),
      panel.grid = element_blank(),
      strip.background = element_blank(),
      legend.position="top",
      legend.title = element_blank(),
      strip.text = element_text(size = 13)
    )
}


```


## Figure labels
```{r}
site_labels <- c(
  "DISHC" = "Clay Loam",
  "DISHT" = "Loam",
  "DISHR" = "Sandy Loam"
)

```

# Import and Combine Data 

## Import data
```{r}
disturbance_fluxes <- 
  data_file %>% 
  read_xlsx(sheet = "OxicIncubation")

anoxic_dist_flux <- 
  data_file %>% 
  read_xlsx(sheet = "AnoxicIncubation")

ox_en_fluxes <- 
  data_file %>% 
  read_xlsx(sheet = "OxygenEnrichment") %>% 
  mutate(
    across("site", factor, levels = c("DISHC", "DISHT", "DISHR")),
    across("wfps_bin", factor, levels = c("15 - 60 %", "> 60 %"))
  )

dist_grav <- 
  data_file %>% 
  read_xlsx(sheet = "GravimetricData - Oxic")

dist_mass_time <- 
  data_file %>% 
  read_xlsx(sheet = "MassOverTime") %>% 
  #Make the pre-disturbance weights the same as the Day 0 After Disturbance (i.e., before subsampling)
  bind_rows(
    data_file %>% 
      read_xlsx(sheet = "MassOverTime") %>% 
      filter(day == 0) %>% 
      select(-c(day, timing)) %>% 
      mutate(timing = "Before Disturbance") %>% 
      add_column(day = 7) %>% 
      add_row(site = "DISHC", core = c(1:25), timing = "Before Disturbance", day = 1) %>% 
      add_row(site = "DISHC", core = c(1:25), timing = "Before Disturbance", day = 2) %>% 
      add_row(site = "DISHC", core = c(1:25), timing = "Before Disturbance", day = 3) %>% 
      add_row(site = "DISHC", core = c(1:25), timing = "Before Disturbance", day = 4) %>% 
      add_row(site = "DISHT", core = c(1:25), timing = "Before Disturbance", day = 1) %>% 
      add_row(site = "DISHT", core = c(1:25), timing = "Before Disturbance", day = 2) %>% 
      add_row(site = "DISHT", core = c(1:25), timing = "Before Disturbance", day = 3) %>% 
      add_row(site = "DISHT", core = c(1:25), timing = "Before Disturbance", day = 4) %>% 
      add_row(site = "DISHR", core = c(1:25), timing = "Before Disturbance", day = 1) %>% 
      add_row(site = "DISHR", core = c(1:25), timing = "Before Disturbance", day = 2) %>% 
      add_row(site = "DISHR", core = c(1:25), timing = "Before Disturbance", day = 3) %>% 
      add_row(site = "DISHR", core = c(1:25), timing = "Before Disturbance", day = 4) %>% 
      arrange(site, core, day) %>% 
      fill(dry_mass, .direction = "up") %>% 
      select(site, core, day, timing, dry_mass)
  ) %>% 
  mutate(across(timing, factor, levels = c("Before Disturbance", "After Disturbance")))

tc_tn <- 
  data_file %>% 
  read_xlsx(sheet = "TotalCarbon")


```

## Combine Data 

### Disturbance Incubation - Oxic 

This code chunk:

* joins the flux data with the average percent C and percent N for the corresponding site
* joins the flux data with the soil dry mass
* joins the flux data with the water content, volume, and bulk density data 
* filters out any cores that were below 15% WFPS
* turns site, wfps, and disturbance into factor variables for ordering
* calculates g of soil C for each observation
* calculates new flux unit

```{r}

oxic_dist_data <- 
  disturbance_fluxes %>% 
  left_join(
    tc_tn %>% 
      group_by(site) %>% 
      summarise(
        avg_perc_c = mean(perc_c, na.rm = TRUE),
        avg_perc_n = mean(perc_n, na.rm = TRUE),
        se_perc_c = sd(perc_c, na.rm = TRUE) / sqrt(n()),
        se_perc_n = sd(perc_n, na.rm = TRUE) / sqrt(n())
      ), 
    by = "site"
  ) %>% 
  left_join(
    dist_mass_time, 
    by = c("site", "timing", "core_num" = "core", "day")
  ) %>% 
  left_join(dist_grav, by = c("site", "core_num" = "core")) %>% 
  filter(wfps >= 15) %>% 
  mutate(
    across("site", factor, levels = c("DISHC", "DISHT", "DISHR")),
    across(c("timing", "control_disturbed"), factor),
    wfps_bin = 
      factor(
        if_else(wfps <= 60, "15 - 60 %", "> 60 %"), 
        levels = c("15 - 60 %", "> 60 %")
      ),
    g_soil_c = dry_mass * avg_perc_c / 100,
    flux_umol_g_soil_C_hr = flux * soil_core_area_m2 * s_in_hr / g_soil_c
  ) 

```


### Disturbance Incubation - Anoxic Control

This code chunk:

* joins the anoxic flux data with carbon data 
* joins the anoxic flux data with the average bulk density of each core 
* calculates g of soil C in each core
* calculates new flux unit

```{r}
anoxic_dist_data <-
  anoxic_dist_flux %>% 
  left_join(
    tc_tn %>% 
      group_by(site) %>% 
      summarise(
        avg_perc_c = mean(perc_c, na.rm = TRUE),
        avg_perc_n = mean(perc_n, na.rm = TRUE),
        se_perc_c = sd(perc_c, na.rm = TRUE) / sqrt(n()),
        se_perc_n = sd(perc_n, na.rm = TRUE) / sqrt(n())
      ), 
    by = "site"
  ) %>% 
  filter(wfps >= 15) %>% 
  mutate(
    across("site", factor, levels = c("DISHC", "DISHT", "DISHR")),
    g_soil_c = dry_wt * avg_perc_c / 100,
    co2_umol_g_soil_C_hr = 
      co2_flux_umol_m2_s * anoxic_soil_core_area_m2 * s_in_hr / g_soil_c,
    ch4_umol_g_soil_C_hr = 
      ch4_flux_umol_m2_s * anoxic_soil_core_area_m2 * s_in_hr / g_soil_c,
    n2o_umol_g_soil_C_hr = 
      ((n2o_accum / 10^6) * 10^6 * jar_volume_l * mols_per_l) / 
      (g_soil_c * elapsed_time_hr),
    wfps_bin = 
      factor(
        if_else(wfps <= 60, "15 - 60 %", "> 60 %"), 
        levels = c("15 - 60 %", "> 60 %")
      ),
  ) 


```

# Figures and Analysis

## Physical disturbance incubation (oxic)

### Figure 2: Efflux from physical disturbance incubation 

#### Isolate Before Disturbance Day 7 Fluxes (at WFPS bin level)

```{r}
bd_7_flux_umol_CO2_gC_hr <-
  oxic_dist_data %>% 
  filter(timing == "Before Disturbance", day == 7, wfps_bin == "15 - 60 %") %>%          
  filter(!(core_num %in% control_cores)) %>% 
  group_by(site) %>% 
  summarise(
    bd_7_avg_flux = mean(flux_umol_g_soil_C_hr, na.rm = TRUE),
    bd_7_se_flux = sd(flux_umol_g_soil_C_hr, na.rm = TRUE)/sqrt(n())
  ) 


```

#### Isolate Before Disturbance Day 7 Fluxes (at Core level)

```{r}
bd_7_flux_umol_CO2_gC_hr_core <-
  oxic_dist_data %>% 
  filter(timing == "Before Disturbance", day == 7) %>% 
  group_by(site, wfps_bin, core_num) %>% 
  summarise(
    bd_7_avg_flux = mean(flux, na.rm = TRUE),
    bd_7_se_flux = sd(flux, na.rm = TRUE)/sqrt(n())
  ) 

```

#### Make figure

```{r}
oxic_dist_data %>% 
  filter(wfps_bin == "15 - 60 %" & !(core_num %in% control_cores)) %>% 
  left_join(bd_7_flux_umol_CO2_gC_hr, by = c("site")) %>% 
  group_by(
    site, 
    wfps_bin, 
    timing, 
    day, 
    cumulative_day, 
    bd_7_avg_flux, 
    bd_7_se_flux, 
    control_disturbed
  ) %>% 
  summarize(
    avg_flux = mean(flux_umol_g_soil_C_hr, na.rm = TRUE),
    se_flux = sd(flux_umol_g_soil_C_hr, na.rm = TRUE) / sqrt(n())
  ) %>% 
  filter(
    control_disturbed == "Undisturbed" & cumulative_day <= 7 | 
      control_disturbed == "Disturbed"
  ) %>% 
  ggplot(
    aes(
      x = cumulative_day, 
      color = control_disturbed, 
      group = control_disturbed
    )
  ) + 
  geom_vline(aes(xintercept = 7), color = "black", linetype = 3) +
  geom_rect(
      xmin = 7,
      xmax = Inf,
    aes(
      ymin = bd_7_avg_flux - bd_7_se_flux, 
      ymax = bd_7_avg_flux + bd_7_se_flux
    ),
    color = "gray",
    fill = "gray",
    alpha = 0.05,
    size = 0.2
  ) + 
  geom_segment(
    aes(y = bd_7_avg_flux, yend = bd_7_avg_flux),
    x = 7, 
    xend = Inf, 
    linetype = 2
  ) +
  geom_pointrange(
    aes(
      y = avg_flux,
      ymin = avg_flux - se_flux,
      ymax = avg_flux + se_flux
    ),
    size = 0.25
  ) + 
  geom_line(aes(y = avg_flux), linetype = 2) +
  scale_x_continuous(breaks = c(1, 7, 14, 21)) +
  scale_color_manual(
    values = c("red", "black"),
    guide = guide_legend(reverse = TRUE)
  ) +
  annotate(
    "rect", 
    xmin = -Inf, 
    xmax = 7, 
    alpha = .15, 
    ymin = -Inf, 
    ymax = Inf
  ) +
  facet_grid(
    cols = vars(site), 
    labeller = labeller(site = site_labels)
  ) +
  my_theme() +
  theme(
    aspect.ratio = 1, 
    legend.text = element_text(size = 12), 
    legend.position = "top",
    panel.background = element_rect(fill = "transparent",colour = NA), 
    plot.background = element_rect(fill = "transparent",colour = NA)
  ) +
  labs(
    x = "Time Since Wetting (days)",
    y = expression(Efflux~"("~μmol~CO[2]~g~soil~C^-1~h^-1~")")
  )


```

#### Table S1 - Reps per site: disturbed 
```{r}
oxic_dist_data %>% 
  filter(wfps_bin == "15 - 60 %" & !(core_num %in% control_cores)) %>% 
  left_join(bd_7_flux_umol_CO2_gC_hr, by = c("site")) %>% 
  group_by(
    site, 
    wfps_bin, 
    timing, 
    day, 
    cumulative_day, 
    bd_7_avg_flux, 
    bd_7_se_flux, 
    control_disturbed
  ) %>% 
  summarize(
    avg_flux = mean(flux_umol_g_soil_C_hr, na.rm = TRUE),
    se_flux = sd(flux_umol_g_soil_C_hr, na.rm = TRUE) / sqrt(n()),
    reps = n()
  ) %>% 
  filter(cumulative_day == 21)

```


#### Table S1 - Reps per site: Undisturbed (after disturbance)
```{r}

oxic_dist_data %>% 
  filter(core_num %in% control_cores) %>% 
  left_join(bd_7_flux_umol_CO2_gC_hr, by = c("site")) %>% 
  group_by(
    site, 
    wfps_bin, 
    timing, 
    day, 
    cumulative_day, 
    bd_7_avg_flux, 
    bd_7_se_flux, 
    control_disturbed
  ) %>% 
  summarize(
    avg_flux = mean(flux_umol_g_soil_C_hr, na.rm = TRUE),
    se_flux = sd(flux_umol_g_soil_C_hr, na.rm = TRUE) / sqrt(n()),
    reps = n()
  ) %>% 
  filter(cumulative_day == 17)

```


#### Inset values - CO2 evolved over 14 days 

The code below 

* conservatively assumes that two cores that were not measured the same day as disturbance treatment had no difference in CO2 for that day

```{r}
oxic_dist_cumulative_flux <- 
  oxic_dist_data %>% 
  filter(timing == "After Disturbance") %>%
  filter(wfps_bin == "15 - 60 %" & !(core_num %in% control_cores)) %>% 
  group_by(site, core_num, timing, cumulative_day, wfps_bin) %>%
  summarise(
    avg_ad_flux = mean(flux_umol_g_soil_C_hr, na.rm = TRUE),
    se_ad_flux = sd(flux_umol_g_soil_C_hr, na.rm = TRUE)/sqrt(n())
  ) %>%
  left_join(
    bd_7_flux_umol_CO2_gC_hr_core, 
    by = c("site", "wfps_bin", "core_num")
  ) %>% 
  pivot_wider(
    id_cols = c(
      site,
      core_num,
      wfps_bin,
      bd_7_avg_flux,
      bd_7_se_flux
    ),
    names_from = cumulative_day,
    values_from = avg_ad_flux
  ) %>% 
  rename_with(
    ~str_c("adflux_",.), 
    .cols = c(`7`, `8`, `9`, `10`, `14`, `17`, `21`)
  ) %>% 
  filter(!core_num %in% control_cores) %>% 
  mutate(
    across(adflux_7, ~if_else(is.na(.), bd_7_avg_flux, .)),
    trap1 = 1 * hr_in_day * (adflux_7  + adflux_8) / 2,
    trap2 = 1 * hr_in_day * (adflux_8 + adflux_9) / 2,
    trap3 = 1 * hr_in_day * (adflux_9 + adflux_10) / 2,
    trap4 = 4 * hr_in_day * (adflux_10 + adflux_14) / 2,
    trap5 = 3 * hr_in_day * (adflux_14 + adflux_17)/ 2,
    trap6 = 4 * hr_in_day * (adflux_17 + adflux_21) / 2,
    total_co2_mmol = 
      (trap1 + trap2 + trap3 + trap4 + trap5 + trap6) * umol_to_mmol,
    hypothetical_undist_co2_min = 14 * bd_7_avg_flux * hr_in_day * umol_to_mmol
  )
```


```{r}
oxic_dist_cumulative_flux %>% 
  group_by(site, wfps_bin) %>% 
  summarize(
    mean_mmolCO2_g_soilC = mean(total_co2_mmol, na.rm = TRUE),
    se_mmolCO2_g_soilC = sd(total_co2_mmol, na.rm = TRUE) / sqrt(n()),
    mean_hypothetical_undist_co2 = 
      mean(hypothetical_undist_co2_min, na.rm = TRUE),
    se_hypothetical_undist_co2 = 
      sd(hypothetical_undist_co2_min, na.rm = TRUE) / sqrt(n()),
  ) %>% 
  pivot_longer(
    cols = contains("co2"),
    names_to = "data_type",
    values_to = "flux"
  ) %>% 
  separate(data_type, c("data_type","dist_hypothetical","drop1","drop2")) %>% 
  select(-c(drop1, drop2)) %>% 
  mutate(
    across(dist_hypothetical, ~if_else(. == "hypothetical", ., "dist"))
  ) %>% 
  pivot_wider(
    id_cols = c(site, wfps_bin, dist_hypothetical),
    names_from = data_type,
    values_from = flux
  ) %>%
  mutate(
    across(
      dist_hypothetical, 
      factor, 
      levels = c("hypothetical", "dist"),
      labels = c("Calculated", "Disturbed")
    )
  )
```


### Figure S1 - Disturbance incubation control cores 

NOTE - DISHR core is at 66% WFPS; this is noted in the figure caption.
```{r}
oxic_dist_data %>% 
  group_by(
    site,
    wfps_bin,
    timing,
    control_disturbed,
    core_num,
    day,
    cumulative_day
  ) %>% 
  filter(
    control_disturbed == "Undisturbed", 
    core_num %in% control_cores,
    (wfps_bin == "15 - 60 %" | (core_num == 9 & site == "DISHR"))
  ) %>% 
  summarise(
    avg_flux = mean(flux_umol_g_soil_C_hr, na.rm = TRUE),
    se_flux = sd(flux_umol_g_soil_C_hr, na.rm = TRUE) / sqrt(n())
  ) %>% 
  ggplot(
    aes(
      x = cumulative_day,
      y = avg_flux,
      group = core_num
    )
  ) +
  geom_line() + 
  geom_vline(aes(xintercept = 7), color = "black", linetype = 2) +
  geom_pointrange(
    aes(
      ymin = avg_flux - se_flux,
      ymax = avg_flux + se_flux
    ),
    color = "grey11",
    fatten = 2
  ) + 
  scale_x_continuous(breaks = c(1, 7,  14, 21)) +
  annotate("rect", xmin = -Inf, xmax = 7, alpha = .15, ymin = -Inf, ymax = Inf) +
  my_theme() +
  theme(aspect.ratio = 1) +
  facet_grid(
    cols = vars(site),
    scales = "free_y",
    labeller = labeller(site = site_labels)
  ) +
  labs(
    x = "Time since moisture adjustment (days)",
    y = expression(Efflux~"("~μmol~CO[2]~g~soil~C^-1~h^-1~")")
  )


```

## Physical disturbance incubation (anoxic)

### Isolate undisturbed and pre-disturbance fluxes

```{r}
anoxic_control_cores <- 
  anoxic_dist_data %>% 
  filter(control_disturbed == "Undisturbed", timing == "AD", day == 14) %>% 
  pull(core)

```

```{r}

bd_7_anoxic_flux <-
  anoxic_dist_data %>% 
  filter(!core %in% anoxic_control_cores) %>% 
  filter(timing == "BD", day == 7) %>% 
  select(site, core, wfps_bin, co2_umol_g_soil_C_hr, ch4_umol_g_soil_C_hr) %>% 
  group_by(site, wfps_bin) %>% 
  summarise(
    avg_bd_7_co2_flux = mean(co2_umol_g_soil_C_hr, na.rm = TRUE),
    avg_bd_7_ch4_flux = mean(ch4_umol_g_soil_C_hr, na.rm = TRUE),
    se_bd_7_co2_flux = sd(co2_umol_g_soil_C_hr, na.rm = TRUE) / sqrt(n()),
    se_bd_7_ch4_flux = sd(ch4_umol_g_soil_C_hr, na.rm = TRUE) / sqrt(n()),
  )


```


```{r}
bd_7_anoxic_flux_core <-
  anoxic_dist_data %>% 
  filter(!core %in% anoxic_control_cores) %>% 
  filter(timing == "BD", day == 7) %>% 
  select(site, core, wfps_bin, co2_umol_g_soil_C_hr, ch4_umol_g_soil_C_hr) %>% 
  rename(
    bd_7_co2_flux = co2_umol_g_soil_C_hr,
    bd_7_ch4_flux = ch4_umol_g_soil_C_hr
  )

```

### Figure 3 - Anoxic CO2 efflux

```{r}
anoxic_dist_data %>% 
  filter(wfps_bin == "15 - 60 %") %>% 
  group_by(site, wfps_bin, control_disturbed, cumulative_day) %>% 
  summarize(
    avg_co2 = mean(co2_umol_g_soil_C_hr, na.rm = TRUE),
    se_co2 = sd(co2_umol_g_soil_C_hr, na.rm = TRUE) / sqrt(n()),
    avg_ch4 = mean(ch4_umol_g_soil_C_hr, na.rm = TRUE),
    se_ch4 = sd(ch4_umol_g_soil_C_hr) / sqrt(n())
  ) %>% 
  left_join(bd_7_anoxic_flux, by = c("site", "wfps_bin")) %>% 
  filter(
    control_disturbed == "Disturbed"|
      control_disturbed == "Undisturbed" & cumulative_day < 8
  ) %>% 
  ggplot(
    aes(
      x = cumulative_day,
      y = avg_co2,
      color = control_disturbed
    )
  ) +
  geom_rect(
    aes(
      xmin = 7,
      xmax = Inf,
      ymin = avg_bd_7_co2_flux - se_bd_7_co2_flux, 
      ymax = avg_bd_7_co2_flux + se_bd_7_co2_flux
    ),
    color = "gray",
    fill = "gray",
    alpha = 0.05,
    size = 0.2
  ) + 
  geom_vline(aes(xintercept = 7), color = "black", linetype = 3) +
  geom_segment(
    aes(
      y = avg_bd_7_co2_flux, 
      yend = avg_bd_7_co2_flux,
      x = 7,
      xend = Inf
      ),
    linetype = 2
  ) + 
  geom_line() + 
  geom_pointrange(
    aes(
      ymin = avg_co2 - se_co2,
      ymax = avg_co2 + se_co2
    ),
    fatten = 2.5
  ) + 
  scale_x_continuous(breaks = c(1, 7, 14, 21)) +
  scale_color_manual(
    values = c("red","black"),
    guide = guide_legend(reverse = TRUE)
  ) +
  annotate("rect", xmin = -Inf, xmax = 7, alpha = .15, ymin = -Inf, ymax = Inf) +
  my_theme() +
  theme(
    legend.position = "top", 
    aspect.ratio = 1,
    legend.text = element_text(size = 12)
  ) +
  facet_grid(
    rows = vars(wfps_bin), 
    cols = vars(site),
    scales = "free_y",
    labeller = labeller(site = site_labels)
  ) +
  labs(
    x = "Time since moisture adjustment (days)",
    y = expression(Efflux~"("~μmol~CO[2]~g~soil~C^-1~h^-1~")")
  )


```

### Table S1 - Calculating replicates

```{r}
anoxic_dist_data %>% 
  filter(wfps_bin == "15 - 60 %") %>% 
  group_by(site, wfps_bin, control_disturbed, cumulative_day) %>% 
  summarize(
    avg_co2 = mean(co2_umol_g_soil_C_hr, na.rm = TRUE),
    se_co2 = sd(co2_umol_g_soil_C_hr, na.rm = TRUE) / sqrt(n()),
    avg_ch4 = mean(ch4_umol_g_soil_C_hr, na.rm = TRUE),
    se_ch4 = sd(ch4_umol_g_soil_C_hr) / sqrt(n()),
    reps = n()
  ) %>% 
  filter(cumulative_day == 17) %>% 
  select(site, wfps_bin, control_disturbed, reps)
```


### Figure S2 - Anoxic disturbance incubation

```{r}
anoxic_dist_data %>% 
  filter(wfps_bin == "15 - 60 %") %>% 
  group_by(site, wfps_bin, control_disturbed, cumulative_day) %>% 
  summarize(
    avg_co2 = mean(co2_umol_g_soil_C_hr, na.rm = TRUE),
    se_co2 = sd(co2_umol_g_soil_C_hr, na.rm = TRUE) / sqrt(n()),
    avg_ch4 = mean(ch4_umol_g_soil_C_hr, na.rm = TRUE),
    se_ch4 = sd(ch4_umol_g_soil_C_hr) / sqrt(n())
  ) %>% 
  left_join(bd_7_anoxic_flux, by = c("site", "wfps_bin")) %>% 
  ggplot(
    aes(
      x = cumulative_day,
      y = avg_co2,
      color = control_disturbed
    )
  ) +
  geom_rect(
    aes(
      xmin = 7,
      xmax = Inf,
      ymin = avg_bd_7_co2_flux - se_bd_7_co2_flux, 
      ymax = avg_bd_7_co2_flux + se_bd_7_co2_flux
    ),
    color = "gray",
    fill = "gray",
    alpha = 0.05,
    size = 0.2
  ) + 
  geom_vline(aes(xintercept = 7), color = "black", linetype = 3) +
  geom_segment(
    aes(
      y = avg_bd_7_co2_flux, 
      yend = avg_bd_7_co2_flux,
      x = 7,
      xend = Inf
      ),
    linetype = 2
  ) + 
  geom_line() + 
  geom_pointrange(
    aes(
      ymin = avg_co2 - se_co2,
      ymax = avg_co2 + se_co2
    ),
    fatten = 2.5
  ) + 
  scale_x_continuous(breaks = c(1, 7, 14, 21)) +
  scale_color_manual(
    values = c("red","black"),
    guide = guide_legend(reverse = TRUE)
  ) +
  annotate("rect", xmin = -Inf, xmax = 7, alpha = .15, ymin = -Inf, ymax = Inf) +
  my_theme() +
  theme(
    legend.position = "top", 
    aspect.ratio = 1,
    legend.text = element_text(size = 12)
  ) +
  facet_grid(
    rows = vars(wfps_bin), 
    cols = vars(site),
    scales = "free_y",
    labeller = labeller(site = site_labels)
  ) +
  labs(
    x = "Time since moisture adjustment (days)",
    y = expression(Efflux~"("~μmol~CO[2]~g~soil~C^-1~h^-1~")")
  )


```


### Figure 3 Inset: Excess CO2 Evolved in Disturbed vs Calculated (Anoxic disturbance)

```{r}
anoxic_dist_data %>% 
  filter(!core %in% anoxic_control_cores) %>% 
  filter(timing == "AD") %>% 
  pivot_wider(
    id_cols = c(site, core, wfps_bin),
    names_from = cumulative_day,
    names_prefix = "adflux_",
    values_from = co2_umol_g_soil_C_hr
  ) %>% 
  left_join(bd_7_anoxic_flux_core, by = c("site", "core", "wfps_bin")) %>%
  filter(wfps_bin == "15 - 60 %") %>% 
  mutate(
    across(adflux_17, ~if_else(is.na(.), (adflux_14 + adflux_21) / 2, .)),
    trap1 = 1 * hr_in_day * (adflux_7  + adflux_8) / 2,
    trap2 = 1 * hr_in_day * (adflux_8 + adflux_9) / 2,
    trap3 = 1 * hr_in_day * (adflux_9 + adflux_10) / 2,
    trap4 = 4 * hr_in_day * (adflux_10 + adflux_14) / 2,
    trap5 = 3 * hr_in_day * (adflux_14 + adflux_17)/ 2,
    trap6 = 4 * hr_in_day * (adflux_17 + adflux_21) / 2,
    total_co2_mmol = 
      (trap1 + trap2 + trap3 + trap4 + trap5 + trap6) * umol_to_mmol,
    hypothetical_undist_co2_min = 14 * bd_7_co2_flux * hr_in_day * umol_to_mmol
  ) %>% 
  group_by(site, wfps_bin) %>% 
  summarize(
    mean_mmolCO2_g_soilC = mean(total_co2_mmol, na.rm = TRUE),
    se_mmolCO2_g_soilC = sd(total_co2_mmol, na.rm = TRUE)/sqrt(n()),
    mean_hypothetical_undist_co2 = 
      mean(hypothetical_undist_co2_min, na.rm = TRUE),
    se_hypothetical_undist_co2 = 
      sd(hypothetical_undist_co2_min, na.rm = TRUE)/sqrt(n()),
  ) %>% 
  pivot_longer(
    cols = contains("co2"),
    names_to = "data_type",
    values_to = "flux"
  ) %>% 
  separate(data_type, c("data_type","dist_hypothetical","drop1","drop2")) %>% 
  select(-c(drop1, drop2)) %>% 
  mutate(
    across(dist_hypothetical, ~if_else(. == "hypothetical", ., "dist"))
  ) %>% 
  pivot_wider(
    id_cols = c(site, wfps_bin, dist_hypothetical),
    names_from = data_type,
    values_from = flux
  ) %>%
  mutate(
    across(
      dist_hypothetical, 
      factor, 
      levels = c("hypothetical", "dist"),
      labels = c("Calculated", "Disturbed")
    )
  )
```

## Oxygen Enrichment 

### Figure 1 - Oxygen enrichment flux

```{r warning=FALSE}

ox_en_fluxes %>% 
  mutate(
    across(c(ch4_flux_umol_m2_s_norm, n2o_flux_umol_m2_s), as.numeric),
    co2_flux_umol_hr_g_soilC = 
      co2_flux_umol_m2_s_norm * soil_core_area_m2 * s_in_hr,
    ch4_flux_umol_hr_g_soilC = 
      ch4_flux_umol_m2_s_norm * soil_core_area_m2 * s_in_hr,
    n2o_flux_umol_hr = 
      n2o_flux_umol_m2_s * soil_core_area_m2 * s_in_hr
  ) %>% 
  filter(site %in% c("DISHC", "DISHT", "DISHR")) %>% 
  group_by(
    site,
    ox_enrich,
    wfps_bin,
    timing,
    day,
    cumulative_day
  ) %>% 
  summarise(
    avg_co2_umol_hr_g_C = mean(co2_flux_umol_hr_g_soilC, na.rm = TRUE),
    avg_ch4_umol_hr_g_C = mean(ch4_flux_umol_hr_g_soilC, na.rm = TRUE),
    avg_ch4_umol_hr = mean(n2o_flux_umol_hr, na.rm = TRUE),
    se_co2_umol_hr_g_C = sd(co2_flux_umol_hr_g_soilC, na.rm = TRUE) / sqrt(n()),
    se_ch4_umol_hr_g_C = sd(ch4_flux_umol_hr_g_soilC, na.rm = TRUE) / sqrt(n()),
    se_n2o_umol_hr = sd(n2o_flux_umol_hr, na.rm = TRUE) / sqrt(n())
  ) %>% 
  ggplot(
    aes(x = cumulative_day, y = avg_co2_umol_hr_g_C, color = ox_enrich)
  ) + 
  geom_vline(aes(xintercept = 7), color = "black", linetype = 3) +
  geom_pointrange(
    aes(
      ymin = avg_co2_umol_hr_g_C - se_co2_umol_hr_g_C,
      ymax = avg_co2_umol_hr_g_C + se_co2_umol_hr_g_C
    ),
    size = 0.25
  ) + 
  geom_line() + 
  annotate("rect", xmin = -Inf, xmax = 7, alpha = .15, ymin = -Inf, ymax = Inf) +
  scale_color_manual(values = c("black", "red")) +
  scale_x_continuous(breaks = c(1, 7, 14, 21)) + 
  scale_y_continuous(breaks = c(0, 4, 8, 12)) +
  facet_grid(
    cols = vars(site), 
    rows = vars(wfps_bin), 
    labeller = labeller(site = site_labels)
  ) +
  my_theme() +
  theme(
    aspect.ratio = 1, 
    legend.text = element_text(size = 12), 
    legend.position = "top",
    panel.background = element_rect(fill = "transparent",colour = NA), 
    plot.background = element_rect(fill = "transparent",colour = NA),
  ) +
  labs(
    x = "Time Since Wetting (days)",
    y = expression(Efflux~"("~μmol~CO[2]~g~soil~C^-1~h^-1~")")
  )
  
  
```

### Statistics: atmospheric vs enriched efflux by day

#### Check normality

```{r warning=FALSE}
day_by_day <- 
ox_en_fluxes %>% 
  mutate(
    across(c(ch4_flux_umol_m2_s_norm, n2o_flux_umol_m2_s), as.numeric),
    co2_flux_umol_hr_g_soilC = 
      co2_flux_umol_m2_s_norm * soil_core_area_m2 * s_in_hr,
    ch4_flux_umol_hr_g_soilC = 
      ch4_flux_umol_m2_s_norm * soil_core_area_m2 * s_in_hr,
    n2o_flux_umol_hr = 
      n2o_flux_umol_m2_s * soil_core_area_m2 * s_in_hr
  ) %>% 
  filter(site %in% c("DISHC", "DISHT", "DISHR"))


day_by_day %>% 
  filter(timing == "AD") %>% 
  group_by(site, timing, day, ox_enrich, wfps_bin) %>% 
  summarise(
    normality_p = shapiro.test(co2_flux_umol_hr_g_soilC)$p.value
  ) %>% 
  arrange(normality_p)

```
It appears that the DISHC AD 1 > 60% WFPS is the only non-normal distribution...


```{r}
day_by_day %>% 
  filter(site == "DISHC", timing == "AD", day == 1) %>% 
  ggplot(aes(x = co2_flux_umol_hr_g_soilC)) + 
  geom_histogram() + 
  facet_grid(cols = vars(ox_enrich))

day_by_day %>% 
  filter(timing == "AD") %>% 
  group_by(site, timing, day, ox_enrich, wfps_bin) %>% 
  summarise(
    normality_p = shapiro.test(co2_flux_umol_hr_g_soilC)$p.value
  ) %>% 
  arrange(normality_p)
```


It appears all of the fluxes on each day are normally distributed. 

#### Check for equal variances 

```{r}
leveneTest(
  co2_flux_umol_hr_g_soilC ~ site*timing*day*ox_enrich*wfps_bin, 
  data = day_by_day %>% mutate(across(day, factor))
)

```
It appears that there are equal variances.


#### T-test(s)

For every day that it is valid, I will do a t-test between the Atmospheric and Enriched treatment. 

```{r}

day_by_day %>% 
  filter(!(site == "DISHC" & timing == "AD" & day == 1)) %>% 
  filter(timing == "AD") %>% 
  group_by(site, timing, day, ox_enrich, wfps_bin) %>% 
  summarise(
    normality_p = shapiro.test(co2_flux_umol_hr_g_soilC)$p.value
  ) %>% 
  arrange(normality_p)
```

Marginally significant and significant results


```{r warning=FALSE}

day_by_day %>% 
  filter(!(site == "DISHC" & timing == "AD" & day == 1)) %>% 
  filter(timing == "AD") %>% 
  unite(site, timing, day, wfps_bin, col = category) %>% 
  select(category, ox_enrich, co2_flux_umol_hr_g_soilC) %>% 
  pivot_wider(
    id_cols = category, 
    names_from = ox_enrich, 
    values_from = co2_flux_umol_hr_g_soilC
  ) %>% 
  group_by(category) %>% 
  mutate(
    t_test = map2_dbl(Atmospheric, Enriched, ~t.test(.x, .y)$p.value)
  ) %>% 
  separate(category, into = c("site", "timing", "day", "wfps_bin")) %>% 
  filter(t_test < 0.1) %>% 
  arrange(site, timing, as.integer(day))



```

All results

```{r warning=FALSE}
day_by_day %>% 
  filter(!(site == "DISHC" & timing == "AD" & day == 1)) %>% 
  filter(timing == "AD") %>% 
  unite(site, timing, day, wfps_bin, col = category) %>% 
  select(category, ox_enrich, co2_flux_umol_hr_g_soilC) %>% 
  pivot_wider(
    id_cols = category, 
    names_from = ox_enrich, 
    values_from = co2_flux_umol_hr_g_soilC
  ) %>% 
  group_by(category) %>% 
  mutate(
    t_test = map2_dbl(Atmospheric, Enriched, ~t.test(.x, .y)$p.value)
  ) %>% 
  separate(category, into = c("site", "timing", "day", "wfps_bin")) %>% 
  arrange(site, timing, as.integer(day))
```

## Figure 1 Inset: Cumulative fluxes 

```{r warning=FALSE}
ox_en_co2_tot <- 
  ox_en_fluxes %>% 
  mutate(
    across(c(ch4_flux_umol_m2_s_norm, n2o_flux_umol_m2_s), as.numeric),
    co2_flux_umol_hr_g_soilC = 
      co2_flux_umol_m2_s_norm * soil_core_area_m2 * s_in_hr,
    ch4_flux_umol_hr_g_soilC = 
      ch4_flux_umol_m2_s_norm * soil_core_area_m2 * s_in_hr,
    n2o_flux_umol_hr = 
      n2o_flux_umol_m2_s * soil_core_area_m2 * s_in_hr
  ) %>% 
  filter(site %in% c("DISHC", "DISHT", "DISHR")) %>% 
  filter(timing == "AD" | (timing == "BD" & day == 7)) %>% 
  select(
    site, 
    core_num, 
    wfps_bin,
    cumulative_day, 
    ox_enrich,
    co2_flux_umol_hr_g_soilC,
    ch4_flux_umol_hr_g_soilC,
    n2o_flux_umol_hr
  ) %>% 
  pivot_wider(
    id_cols = c(site, core_num, wfps_bin, ox_enrich), 
    names_from = cumulative_day,
    names_prefix = "adflux_",
    values_from = co2_flux_umol_hr_g_soilC
  ) %>%
  mutate(
    trap1 = 1 * hr_in_day * (adflux_7  + adflux_8) / 2,
    trap2 = 1 * hr_in_day * (adflux_8 + adflux_9) / 2,
    trap3 = 1 * hr_in_day * (adflux_9 + adflux_10) / 2,
    trap4 = 4 * hr_in_day * (adflux_10 + adflux_14) / 2,
    trap5 = 3 * hr_in_day * (adflux_14 + adflux_17)/ 2,
    trap6 = 4 * hr_in_day * (adflux_17 + adflux_21) / 2,
    total_co2_mmol = (trap1 + trap2 + trap3 + trap4 + trap5 + trap6) * umol_to_mmol,
  )

```


```{r}
ox_en_co2_tot %>% 
    group_by(wfps_bin, site, ox_enrich) %>% 
  summarise(
    avg_co2_mmol = mean(total_co2_mmol, na.rm = TRUE),
    se_co2_mmol = sd(total_co2_mmol, na.rm = TRUE) / sqrt(n())
  )
```


### Statistics - Atmospheric vs Enriched 

#### Check normality

```{r}
ox_en_co2_tot %>% 
  group_by(wfps_bin, site, ox_enrich) %>% 
  summarise(
    normality_stat = shapiro.test(total_co2_mmol)$statistic,
    normality_p_value = shapiro.test(total_co2_mmol)$p.value
  ) 
```
All are normally distributed! Yahoo! 

#### Check for equal variances
```{r}

leveneTest(
  total_co2_mmol ~ site*wfps_bin*ox_enrich, 
  data = ox_en_co2_tot 
)
```

Equal variances!

#### T-test(s) 

DISHC - dry

```{r}
t.test(
  total_co2_mmol~ox_enrich, 
  data = ox_en_co2_tot %>% 
  filter(
    site == "DISHC",
    wfps_bin == "15 - 60 %"
  )
)

```


DISHC - wet

```{r}
t.test(
  total_co2_mmol~ox_enrich, 
  data = ox_en_co2_tot %>% 
  filter(
    site == "DISHC",
    wfps_bin == "> 60 %"
  )
)
```


DISHT - dry

```{r}
t.test(
  total_co2_mmol~ox_enrich, 
  data = ox_en_co2_tot %>% 
  filter(
    site == "DISHT",
    wfps_bin == "15 - 60 %"
  )
)
```


DISHT - wet

```{r}
t.test(
  total_co2_mmol~ox_enrich, 
  data = ox_en_co2_tot %>% 
  filter(
    site == "DISHT",
    wfps_bin == "> 60 %"
  )
)
```


DISHR - dry

```{r}
t.test(
  total_co2_mmol~ox_enrich, 
  data =  ox_en_co2_tot %>% 
  filter(
    site == "DISHR",
    wfps_bin == "15 - 60 %"
  )
)
```


DISHR - wet

```{r}
t.test(
  total_co2_mmol~ox_enrich, 
  data = ox_en_co2_tot %>% 
  filter(
    site == "DISHR",
    wfps_bin == "> 60 %"
  )
)
```









